---
title: "HW11"
author: "ASG"
date: "`r Sys.Date()`"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r set up workspace}

setwd("/Users/animeshsengupta/Work Directory/DACSS/STAT625/Homeworks") 
library(MASS)
library(alr4)  # loads the installed package into the workspace so you can use it
library(summarytools)
library(ggplot2)
library(plotly)
library(splines)
library(boot)
library(sandwich)
library(plotly)
```

## Answer 12.1

### Answer 12.1.1
```{r}
colnames(Blowdown)
bd <- as.data.frame.matrix(xtabs(~ spp + y, Blowdown))
bd$total=bd$`0`+bd$`1`
bd$died=bd$`1`
bd
```

### 12.1.2
```{r}
bf<-Blowdown%>%filter(spp=="balsam fir")
ggplot(data=bf, aes(x=log(d), y=y)) + 
        geom_point() +
        geom_smooth()
```
As per the visualization , we can say that the response variable y only admits two values 0 and 1. Additionally plotting the smoother , we can clearly see that it shows a logistic curve. Hence fitting a logistic Generalised linear model would serve best.

### 12.1.3

```{r}
bds <- glm(y ~ log(d), family=binomial,data=Blowdown%>%filter(spp=="black spruce"))
summary(bds)

```
As seen from the summary, the Residual deviance and df are different from 49.891 and  33 df to 655.24 and 657 df.

### 12.1.4
```{r}
bds2 <- glm(y ~ log(d)+I(log(d)^2), family=binomial,data=Blowdown%>%filter(spp=="black spruce"))
summary(bds2)
```
```{r}
anova(bds2,test="Chisq")

plot(Effect("d", bds2), main="")
```
Since z=0.0403 and g=0.044, z^2 and G^2 is definitely not same.
the effect plots doesnt show a declining trend but declining probability cant be discounted, it can be plausible.

## 12.3

### 12.3.1
```{r}
colnames(Downer)
dw<-xtabs(~outcome+myopathy,data=Downer)
dw[2,]/colSums(dw)
```

### 12.3.2
```{r}
dw1<-glm(outcome~myopathy,family=binomial,data=Downer)
summary(dw1)
confint(dw1)
predict(dw1,data.frame(myopathy=factor(levels(Downer$myopathy))),type="response")
```
The intercept is when myopathy = 0, the estimated log-odds of survival . The coefficient for myopathy is the increase in log-odds when myopathy is present.

### 12.3.3
```{r}
dw2 <- glm(outcome~log(ck),family=binomial, data=Downer)
summary(dw2)
```

### 12.3.4
```{r}
dw3 <- glm(outcome~myopathy+log(ck)+myopathy:log(ck),family=binomial,data=Downer)
Anova(dw3)
plot(Effect(c("myopathy", "ck"), dw3),multiline=TRUE)
```

## 12.9
### 12.9.1
```{r}
colnames(AMSsurvey)
AMS1 <- reshape(AMSsurvey, varying=c("count", "count11"), v.names="y",direction="long", times=c("08-09", "11-12"),timevar="year")
AMS1$type <- factor(AMS1$type, levels=levels(AMS1$type)[order(xtabs(y~type, AMS1))])
AMS1$year <- factor(AMS1$year)

ams1<-glm(y ~ (sex + citizen)*type, poisson, AMS1)
anova(ams1,test = "Chisq")
```
#12.9.2
```{r}
ams2 <- glm(y ~ (type + sex + citizen + year)^4, poisson, AMS1)
anova(ams2)
ams3 <- update(ams2, ~ type*(sex + citizen + year))
Anova(ams3)
```

### 12.7.1
```{r}
w1 <- glm(cbind(surv, m - surv) ~ class + age + sex,family=binomial, data=Whitestar)
summary(w1)
```
Since all women survived apart from women in third class , we need to add an interaction term of sex*class.

### 12.7.2
```{r}
w2 <- update(w1, ~(class + age + sex)^2)
Anova(w2)
```