---
title: "Homework 5"
author: "Animesh Sengupta"
date: "10/11/2022"
output: pdf_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r set up workspace}

setwd("/Users/animeshsengupta/Work Directory/DACSS/STAT625/Homeworks") 
library(alr4)  # loads the installed package into the workspace so you can use it
library(summarytools)
library(ggplot2)
library(plotly)
library(splines)
```


## Answer 5.1

### 5.1.1

The mean function is given as 
$$E(Y|U_{2},...,U_{d})=\beta_{0}+\beta_{2}U_{2}+...+\beta_{j}U{j}$$

From the given equation, since U is a dummy variable equal to 1 in the jth level. So from the above equation if j=1, then there are no levels for U in 1st level to be 1, hence all the levels will have value to be 0. Hence for level1 we can modify the above equation as :
$$E(Y|U_{1})=\beta_{0}+0$$
Since for the jth level all other values of U become 0 other than the jth level we can decompose the mean function to :
$$E(Y|U_{j})=\beta_{0}+\beta_{j}\\U_{j}=1$$
## Answer 5.5

### Answer 5.5.1

$$Y \sim A+B+A:B$$
### 5.5.2
$$\beta_{0}=u_{11}\\
\beta_{1}=u_{21}-u_{11}\\
\beta_{2}=u_{12}-u_{11}\\
\beta_{3}=u_{13}-u_{11}\\
\beta_{4}=u_{11}+u_{22}-u_{21}-u_{12}\\
\beta_{5}=u_{11}+u_{22}-u_{21}-u_{13}\\$$
### 5.5.3
Since we are only fitting for the model Y ~ A+B , we dont have mixed interaction terms between A and B, hence in the above equation we can remove the b4 and b5 since there are no mixed interactions considered in the new model. Hence the B in terms of U can be written as:
$$\beta_{0}=u_{11}\\
\beta_{1}=u_{21}-u_{11}\\
\beta_{2}=u_{12}-u_{11}\\
\beta_{3}=u_{13}-u_{11}\\$$

## 5.8

### 5.8.1
```{r}
colnames(cakes)
cl<-lm(Y~X1+X2+I(X1^2)+I(X2^2)+X1:X2,data=cakes)
summary(cl)
```
As per the model summary , we can verify that the significance level of all polynomials terms are  less than 0.005.

### 5.8.2
```{r}
colnames(cakes)
cl<-lm(Y~X1+X2+I(X1^2)+I(X2^2)+X1:X2+block,data=cakes)
summary(cl)
```
When adding the block term in our linear model , we get an interaction term of block1 in our mean function with an estimate. Although the significance level of that estiamte is very high i.e 0.65. Relatively in the previous model and this model , the estimate values hasnt changed and remains the same in both the models.

## Answer 5.9

### 5.9.1

```{r}
colnames(salarygov)
plot(salarygov$Score,salarygov$MaxSalary)
```

As visualised by the plot, a simple linear regression might not be enough to predict accurately the score given the maxsalary because we can see an upwards curve trend. Also we can see a lot of variability at right while a lot many points are consolidated at right.

### 5.9.2
```{r}
sgl0=lm(MaxSalary~Score,salarygov)
sgl1=lm(MaxSalary~bs(Score,4),salarygov)
sgl2=lm(MaxSalary~bs(Score,5),salarygov)
sgl3=lm(MaxSalary~bs(Score,10),salarygov)
plot(salarygov$Score,salarygov$MaxSalary,pch=".")

abline(sgl1,col = "red")
abline(sgl2,col = "green")
abline(sgl3,col = "blue")
```
## 5.11

# 5.11.1
```{r}
colnames(MinnLand)
ml1<-lm(log(acrePrice)~year+region+year:region+financing,data=MinnLand)
summary(ml1)
confint(ml1)["financingseller_financed",]
```

### 5.11.2
1. Seller financing lowers sale prices : Since the estimate of the financingseller_financed is negative as per the aboce summary, we can say that it may imply a certain causation. Although it doesnt mean that seller financing is the only reason to lower the sale price and is the only determinant for lower sale. 

2. Seller financing is more likely on lower-priced property
transactions: Since the model assigned a negative estiimate for financingseller_financed then it means that it has seen observation where the lower priced property is seller financed, but since there are other estimators and terms involved it may not be the only sole reason. 

## 5.19.

### 5.19.1
Wew have to assume that i belongs to (len,amp,load) and dummy variable $$U_{ij}$$ for 3 levels of j i.e 1,2,3
hence we can write first order of mean function as:  
$$E(log(cycles)|First-order)=\beta_0+\sum\limits_{i=1}^3\sum\limits_{j=1}^3\beta_{ij}U_{ij}\\$$
$$E(log(cycles)|Second-order)=\beta_0+\sum\limits_{i=1}^3\sum\limits_{j=1}^3\beta_{ij}U_{ij}+\sum\limits_{i=1}^2\sum\limits_{j=1}^3\sum\limits_{k=i+1}^3\beta_{ijk}U_{ij}U_{kj}$$

### 5.19.2
If the load is increased from middle to higher level , it means that the jth factor increases from 1 to 2 for i=3. This means that the differnce in response will be estimator of i,j=3,3 minus i,j=3,2 i.e
$$ \beta_{33}-\beta_{32} $$

for the second order similarly we will compute the change in interaction terms when i=1 and 2 while changing the level j from 2 to 3 i.e

$$ \beta_{33}-\beta_{32} + \beta_{133}-\beta_{132}+\beta_{233}-\beta_{232} $$