---
title: "Homework 8"
author: "Animesh Sengupta"
date: "10/31/2022"
output: pdf_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r set up workspace}

setwd("/Users/animeshsengupta/Work Directory/DACSS/STAT625/Homeworks") 
library(MASS)
library(alr4)  # loads the installed package into the workspace so you can use it
library(summarytools)
library(ggplot2)
library(plotly)
library(splines)
library(boot)
library(sandwich)
library(plotly)
```

## 8.4

### Answer 8.4.1
```{r}
sglm=lm(log(MaxSalary)~Score,data = salarygov)
summary(sglm)
residualPlots(sglm)
ncvTest(sglm)
```
The residual plot shows signs of a scattered residuals, there are no negative trends which can say that the mean function isnt fitting improperly. It resembles characteristics from a null plot.

## Answer 8.5

### Answer 8.5.1
```{r}
colnames(BigMac2003)
plot_ly(y=~BigMac,x=~FoodIndex,data=BigMac2003)
BigMac2003%>%arrange(desc(BigMac))%>%head(3)%>%select(BigMac,FoodIndex)
```

It is very clear from the graph that relation between BigMac and foodIndex is nonlinear , hence we need a non-linear mean function or need to transform the variables to have a linear relationship and linear mean function.

### Answer 8.5.2
```{r}
bmlm<-lm(BigMac~FoodIndex,data = BigMac2003)
bmbc<-boxcox(BigMac~FoodIndex,data = BigMac2003)

lambda <- bmbc$x[which.max(bmbc$y)]
BigMac2003$BigMac_transform=(BigMac2003$BigMac^(lambda)-1)/lambda
nbmlm<-lm(BigMac_transform~FoodIndex,data = BigMac2003)
plot_ly(y=~BigMac_transform,x=~FoodIndex,data=BigMac2003)
invResPlot(bmlm)
invResPlot(nbmlm)
```
As seen from the inverse response plot , the lambda generated from boxcox method was successfully able to transform the response variable.

### 8.5.3
```{r}
new_BigMac2003<-BigMac2003%>%filter(BigMac!=185&BigMac!=132)
bmbc<-boxcox(BigMac~FoodIndex,data = new_BigMac2003)
lambda <- bmbc$x[which.max(bmbc$y)]
lambda
new_BigMac2003$BigMac_transform=(new_BigMac2003$BigMac^(lambda)-1)/lambda
nbmlm<-lm(BigMac_transform~FoodIndex,data = new_BigMac2003)
plot_ly(y=~BigMac_transform,x=~FoodIndex,data=new_BigMac2003)
invResPlot(nbmlm)
```

## 9.3

### Answer 9.3.1

```{r}
colnames(pipeline)
plot(Lab~Field,data=pipeline)
```
As visible from the scatter plot, the relationship isnt exactly linear. This can be attributed to the fact that for a few changing x the y remains almost constant. This trends is unfortunately very prevalent in this dataset. 

### Answer9.3.2
```{r}
plm=lm(Lab~Field,data=pipeline)
summary(plm)
residualPlots(plm)
ncvTest(plm)
```
from the estimator of field , we get 1.22 which means that the field values are 1.22 times larger than the lab results. It is not a 1-1 relationship between the two. Also finding the residual plots shows that the variance is constant with variance being concentrated on the left and a megaphone like trend.


### Answer 9.3.3
```{r}
summary(plm)$coef[2, 1:2]

plb<-Boot(plm)
summary(plb)

plw<-lm(Lab~Field,weights=1/Field,data=pipeline)
summary(plw)$coef[2, 1:2]

pld<-deltaMethod(plm,"Field",vcov=hccm)
pld$Estimate
pld$SE
```

## 9.11
```{r}
colnames(fuel2001)
fuel2001$Fuel=fuel2001$FuelC/fuel2001$Pop
fuel2001$Dlic=fuel2001$Drivers/fuel2001$Pop

flm<-lm(Fuel ~ Tax+Dlic+Income+log(Miles),data=fuel2001)
t<-studres(flm)
D<-cooks.distance(flm)
cat("State","t_i","d_i","\n")
cat("Alaska",t["AL"],D["AL"],"\n")
cat("New York",t["NY"],D["NY"],"\n")
cat("Hawaii",t["HI"],D["HI"],"\n")
cat("Wyoming",t["WI"],D["WI"],"\n")
cat("DC",t["DC"],D["DC"])
#head(D,20)
```
the largest outlier from studentized t test is of Wyoming. Since the cooks distance is also small compared to the 4 times mean , we can conclude none of them are outliers.

## 9.19
```{r}
colnames(drugcost)
view(dfSummary(drugcost))
plot(drugcost)
```


