---
title: "Homework 9"
author: "Animesh Sengupta"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r set up workspace}

setwd("/Users/animeshsengupta/Work Directory/DACSS/STAT625/Homeworks") 
library(MASS)
library(alr4)  # loads the installed package into the workspace so you can use it
library(summarytools)
library(ggplot2)
library(plotly)
library(splines)
library(boot)
library(sandwich)
library(plotly)
```

## Answer 11.1

### Answer 11.1.1
the parameter $\gamma$ makes the mean fuction non linear, and in this case it is being multiplied to the predictor $\beta_{ij}$. For every different value of $j \in G$, which essentially means each group has its own predictor and hence its own slope .these indicate that there are a bunch of straight line of mean functions concurrent at $x=\gamma$ for every $\beta_{ij}$.

### Answer 11.1.2
```{r}
colnames(sleep1)
#starting values
slm<-lm(TS ~ log(BodyWt):factor(D), sleep1)
summary(slm)
slm1<-nls(TS~ b0 + b11*((D==1)*(log(BodyWt) - gamma))
                + b12*((D==2)*(log(BodyWt) - gamma))
                + b13*((D==3)*(log(BodyWt) - gamma))
                + b14*((D==4)*(log(BodyWt) - gamma))
                + b15*((D==5)*(log(BodyWt) - gamma)),
    data=sleep1,
    start=list(b0=11,b11=-.28,b12=-.59,b13=-.93,b14=-.64,
               b15=-1.65, gamma=0))
summary(slm1)
```